How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-30 tse-paper/content.tex
corrections to lit survey start

When a build or one of its steps fails, developers typically scavenge
the logs for information related to the source of the error.
This
manual activity is tedious and prone to
error~\cite{santolucito2018statically}.
As a single build log can be
over 50 megabytes large~\cite{beller2017oops}, finding the small chunk
of information which is linked to the actual error, is akin to
searching for a needle in a haystack.
The large numbers of unrelated, but ambiguous
warnings, which many build
logs regularly contain, further compound this problem.
In addition to helping
understand and fix build failures, the information richness of build
logs enables a plethora of other applications: with the help of build
logs, we can better understand and categorize CI failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to do ``predictive CI'' on the build or test outcome
and
duration~\cite{ni2017cost,bisong2017built,haghighatkhah2018test,
machalica2019predictive},
or investigate the role of automated static analysis tools (ASATs,
\cite{beller2016analyzing}) in CI
builds~\cite{zampetti2017open}.

@@

In this section, we survey the emerging field of build log
analysis.
We start with a differentiation to the field of
system log analysis.
% ref to fig 2 so it can be here and above fig 3
Then, we describe the methodology (\Cref{fig:lit-survey})
and results of a systematic mapping
study to
determine the state-of-research on build log analysis following the
guidelines by Petersen et
al.~\cite{petersen2008systematic,petersen2015guidelines}.
We close by summarizing how build logs are commonly
analyzed in literature and discuss the implications of these findings.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system
logs which are produced during runtime of the system.
The first difference is that build logs are produced during runtime
of the build of the system, not during runtime of the system itself.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} depicts two
exemplary lines from such a system log.

The first goal in analyzing system log files is generally to
separate the constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters.
Then, a log parser works on this intermediate representation.
Its output is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}.
This structured log can serve as input to various machine learning and
data mining processes.

@@

\Cref{fig:lit-survey} gives an overview of the literature selection
process.
An initial Google Scholar search for variations of ``build log''
returned more than 2.2 million results on the 15th of March 2020, many
of which stem
from unrelated fields such as construction or wood working.
We refined the search criteria to exclude such obvious
non-related fields and to exclude works on systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results.
% Our replication package documents the full search
% queries~\cite{brandt2020chunk-replication}.
To make handling so many search results feasible, we employed a
three-pass filtering strategy:

First, we filtered articles based on (a) their title and (b) the
contextual information Google Scholar displayed on its search results
overview page.
We included works at large that seemed to bear a resemblance to software
engineering and building software.
Following Kitchenham's protocol~\cite{kitchenham2009systematic},
we excluded
works written in a language unintelligible to the authors
(i.e., not English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

@@

Third, since we had been careful to include rather than exclude borderline
works, a deeper evaluation excluded another 15 articles from the 87
works left after the second filter phase.
Multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the analyzing of build logs.
This excluded another 11 articles.
In sum, we extracted data from 61 articles.

\subsection{Literature Classification}
Having trimmed down the set of articles by 95\%, we investigated the
remaining ones in-depth.
We wanted to characterize how research works with build logs.

Specifically, we were interested in
\begin{itemize}
  \item what information the researchers retrieve from build logs and
  what this information is used for (\textbf{RQ1.1}),
  \item which techniques they to process the build logs,
  in how much detail the techniques are described,
	and whether their implemented tools are published
  (\textbf{RQ1.2}),
  \item what kind of build logs the researchers analyze
	and the origin of these logs (\textbf{RQ1.3}).
\end{itemize}

@@

To support the data extraction we created a template that we filled out
for each of the 61 works.
The template contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping survey with a
pilot:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 56 articles evenly between the two authors
and inspected each article as closely as necessary to fill out the
template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.

@@

\Cref{fig:litsur:use} shows the reason for the
investigated articles to analyze build logs.
The majority of articles (46\%) used the outcome of their build log
analysis to drive further research.
Seo et al.~\cite{seo2014programmers}, for example,
categorized the types of compile errors encountered at
Google.
Other articles (10\%) used the gained insights to give feedback to
developers.
For 8\% of the articles was the retrieved information the input
to a next automatic onward processing step:
articles chose
tests that should be part of a reduced but still effective test
suite~\cite{shi2018evaluating} or filled a data structure with
build information, an information source for further tools aiming
to fix a failing build~\cite{vassallo2018un-break}.

% TODO Moritz: add missing figure
We also identified the type of information which the articles
retrieved from the build logs.
In most cases (51\%), the researchers retrieved a chunk of text.
They scanned build logs for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance, they distinguish failures caused by developers from
failures caused
by infrastructure errors~\cite{lindqvist2019detection} or
determine whether projects use automated static analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.,\ counting
lines, summarization) were only retrieved
in one case each.

@@

Of the 61 articles we inspected, 43 mentioned how they analyzed
build logs.
However, only 16 of these described their method in detail.
\Cref{tab:litsur:techniques} presents the techniques that the articles
described for analyzing build logs.

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they describe
a complex parser (including a tokenizer, lexer, etc.) or refer to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``regular expression'' instead of the tag ``parser.''
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with a manual categorization of build logs and selected
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.
The rather general tags ``analysis'' and ``scan'' in
\Cref{tab:litsur:techniques}, stem---similar to the imprecise
statements about parsing---from the
words authors used to describe their techniques.
%such as
%``automatically analyzed [the build logs]''.

>>>>

When a build or one of its steps fails, developers typically scavenge
the logs for information related to the source of the error.
This
manual activity is tedious and prone to
errors~\cite{santolucito2018statically}.
As a single build log can be
over 50 megabytes large~\cite{beller2017oops}, finding the small chunk
of information which is linked to the actual error, is akin to
searching for a needle in a haystack.
The large numbers of unrelated, but ambiguous
warnings, which many build
logs regularly contain, further compound this problem.
In addition to helping
understand and fix build failures, the information richness of build
logs enables a plethora of other applications: with the help of build
logs, we can better understand and categorize CI failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to do ``predictive CI'' on the build or test outcome
and
duration~\cite{ni2017cost,bisong2017built,haghighatkhah2018test,
machalica2019predictive},
or investigate the role of automated static analysis tools (ASATs)
in CI
builds~\cite{zampetti2017open}.

@@

We performed a systematic mapping study on the emerging field of
build log analysis,
to determine how various researchers analyze build logs.
We start this section by explaining how build log analysis differs
from the more established field of
system log analysis.
% ref to fig 2 so it can be here and above fig 3
Then, we describe the methodology (\Cref{fig:lit-survey})
and results of our systematic mapping
survey following the
guidelines by Petersen et
al.~\cite{petersen2008systematic,petersen2015guidelines}.
% We close by summarizing how build logs are commonly
% analyzed in literature and discuss the implications of these findings.

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system logs.
These system logs are produced during the runtime of a system, while
build logs are produced during the runtime of the build of the system.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} depicts two
exemplary lines from such a system log.

When analyzing system log files, the first goal is generally to
separate the constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters.
The output is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}.
This structured log can serve as input to various machine learning and
data mining processes.

@@

\Cref{fig:lit-survey} gives an overview of the literature selection
process.
An initial Google Scholar search for variations of ``build log''
returned more than 2.2 million results on the 15th of March 2020, many
of which stem
from unrelated fields such as construction or wood working.
We refined the search criteria to exclude such obvious
non-related fields and to exclude works on system
logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results.
% Our replication package documents the full search
% queries~\cite{brandt2020chunk-replication}.
To make handling so many search results feasible, we employed a
three-pass filtering strategy:

First, we filtered articles based on (a) their title and (b) the
contextual information Google Scholar displayed on its search results
page.
We included works at large that seemed to bear a resemblance to software
engineering and building software.
Following Kitchenham's protocol~\cite{kitchenham2009systematic},
we excluded
works written in a language unintelligible to the authors
(i.e., not English or German), which disregarded fewer than 1\% of search
hits.
% 'four' not clear and it gets lost that there are multiple search queries
% therefore I'll cut this part completely
% We then unified
% the results of the four search queries based on the link as the
% identifying element, removing 60 articles that appeared in more than
% one search query.
We removed duplicates based on the title and were
left with 256 articles (16\% of the original search results).

@@

% TODO Caro: check article vs.
work occurrences, balance use
Third, since we had been careful to include rather than exclude borderline
works, a deeper evaluation excluded another 15 articles from the 87
works left after the second filter phase.
Multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the analyzing of build logs.
This excluded another 11 articles.
In sum, we classified 61 articles in detail.

\subsection{Literature Classification}
Having trimmed down the set of articles by 95\%, we investigated the
remaining ones in-depth.
Our aim is to characterize how researchers work with build logs.

Specifically, we were interested in
\begin{itemize}
  \item what information the researchers retrieve from build logs and
  what they use this information for (\textbf{RQ1.1}),
  \item which techniques they employ to process the build logs,
  in how much detail the techniques are described,
	and whether their implemented tools are published
  (\textbf{RQ1.2}),
  \item which kinds of build logs the researchers analyze
	and the origin of these logs (\textbf{RQ1.3}).
\end{itemize}

@@

To support the data extraction we created a template which we filled out
for each of the 61 works.
The template contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''.
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping survey with a
pilot:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 56 articles evenly between the two authors
and inspected each article as closely as necessary to fill out the
template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.

@@

% TODO Moritz: add missing figure
% C: this (to be 4b) is closer to 4a, therefore I moved it up
% and the use/purpose down (to be 4c then)
We also identified the type of information which the articles
retrieved from the build logs.
In most cases (51\%), the researchers retrieved a chunk of text.
They scanned build logs for compilation
errors~\cite{clemencic2014new} or
the duration of tasks within the build~\cite{zhang2016android}.
36\% of the articles classified the build logs into categories.
For instance, they distinguish failures caused by developers from
failures caused
by infrastructure errors~\cite{lindqvist2019detection} or
determine whether projects use automated static analysis
tools~\cite{kavaler2019tool}.
Other types of information (e.g.,\ counting
lines, summarization) were only retrieved
in one case each.

\Cref{fig:litsur:use} shows the reason for the
investigated articles to analyze build logs.
The majority of articles (46\%) used the outcome of their build log
analysis to drive further research.
Seo et al.~\cite{seo2014programmers}, for example,
categorized the types of compile errors encountered at
Google.
Other articles (10\%) used the gained insights to give feedback to
developers.
For 8\% of the articles the retrieved information was the input
to a next automatic onward processing step:
articles chose
tests that should be part of a reduced but still effective test
suite~\cite{shi2018evaluating} or filled a data structure with
build information, an information source for further tools aiming
to fix a failing build~\cite{vassallo2018un-break}.

@@

Of the 61 articles we inspected, 43 mentioned how they analyzed
build logs.
% TODO Caro: check method vs.
technique
However, only 16 of these described their method in detail.
We summarize the techniques that the articles
described for analyzing build logs in \Cref{tab:litsur:techniques}.

% TODO Caro: test \emph{parser}, regex, manually, etc...
The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they describe
a complex parser (including a tokenizer, lexer, etc.) or refer to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``Regular Expression'' instead of the tag ``Parser.''
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with a manual categorization of build logs and selected
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% of the investigated articles
employed machine learning such as natural language
processing or information retrieval techniques.
The rather general tags ``analysis'' and ``scan'' in
\Cref{tab:litsur:techniques}, stem---similar to the imprecise
statements about parsing---from the
words authors used to describe their techniques.
%such as
%``automatically analyzed [the build logs]''.
